---
title: "rct2otrbounds"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rct2otrbounds}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rct2otrbounds)
```

Here we use the EMBARC trial (https://clinicaltrials.gov/ct2/show/NCT01407094) to demonstrate how this package can determine an upper limit on the benefit of using treatment rules instead of a treat everyone approach using clinical trial summary statistics. We focus on the initial phase of the trial, where subjects with early onset depression were randomized to either the antidepressant sertraline or a placebo. We use trial provided by the National Institute of Mental Health Data Archive, but summary data is also publicly available [here](https://clinicaltrials.gov/ct2/show/NCT01407094). There are minor differences in the summaries reported at the previous link and those we calculated from the full data. 

We look at the outcome of the 36-question Hamilton Depression Scale at 8 weeks. Lower scores are beneficial on this scale and it ranges from 0 to 52 points. We first ignore that the outcome is bounded to demonstrate the functions available in this case.

## Benefit Calculation for Unbounded Outcomes

We first look at the upper and lower bounds on the treatment effect heterogeneity (variance of the treatment effect) possible given the means and standard deviations in each arm. Confidence intervals cannot be calculated for unbounded outcomes.

```{r}
treatment_effect_heterogeneity_bound(
  s2_1 = 6.53^2,
  s2_0 = 7.52^2,
  mean_1 = 10.73,
  mean_0 = 11.94,
  bounded_outcome = FALSE,
)


```

Next, we look at the upper bound on the benefit, in terms of reduction in Hamilton Score points, possible given the trial means and standard deviations. Again, no confidence intervals are possible.

```{r}

closed_form_ideal_rule_benefit(
  s2_1 = 6.53^2,
  s2_0 = 7.52^2,
  mean_1 = 10.73,
  mean_0 = 11.94,
  bounded_outcome = FALSE,
)

```

## Benefit Calculation for Bounded Outcomes

We now incorporate the knowledge the the outcome lies within a bounded range. In this case, we can create asympotically conservative confidence intervals, and so we provide the sample size in each arm and the desired confidence level. 

```{r}
treatment_effect_heterogeneity_bound(
  s2_1 = 6.53^2,
  s2_0 = 7.52^2,
  n_1 = 114,
  n_0 = 126,
  mean_1 = 10.73,
  mean_0 = 11.94,
  m = 0,
  M = 52,
  level = 0.95,
  bounded_outcome = TRUE,
  conf.int = TRUE
)


```



```{r}

closed_form_ideal_rule_benefit(
  s2_1 = 6.53^2,
  s2_0 = 7.52^2,
  n_1 = 114,
  n_0 = 126,
  mean_1 = 10.73,
  mean_0 = 11.94,
  m = 0,
  M = 52,
  level = 0.95,
  bounded_outcome = TRUE,
  conf.upper = TRUE
)

```

In this case, the knowledge of the outcome range did not improve our closed form bounds. However, knowledge of the outcome range does allow us to create a tight bound on the treatment rule benefit using the linear programming method.

```{r}

lp_ideal_rule_benefit(
  s2_1 = 6.53^2,
  s2_0 = 7.52^2,
  n_1 = 114,
  n_0 = 126,
  mean_1 = 10.73,
  mean_0 = 11.94,
  m = 0,
  M = 52,
  level = 0.95,
  conf.int = TRUE,
  scale = "lower"
)

```
As we can see, the upper bound with the linear programming method is lower than with the closed form method, however the 95% upper confidence bound is higher. This is due to the fact that both the confidence intervals for both the closed form and linear programming methods are asymptotically conservative as opposed to providing the correct asymptotic coverage. The lower bound on the benefit of the optimal treatment rule is 0, meaning that given the observed trial results it is possible that individualized treatment has no benefit.

## Benefit Calculation for Binary Outcomes

For this example we use fabricated data. Suppose that a study similar to EMBARC measured a binary outcome on whether symptom scores improved by a certain number of points by the end of the study. The sample sizes in each arm are the same as in the EMBARC trial. We can calculate a tight bound on the benefit of a treatment rule and an asymptotically conservative confidence interval as shown below.

```{r}

closed_form_ideal_rule_benefit(
  n_1 = 114,
  n_0 = 126,
  mean_1 = 0.63,
  mean_0 = 0.54,
  level = 0.95,
  binary_outcome = TRUE,
  conf.upper = TRUE
)

```

